% The Acknowledgements page, for thanking everyone
\clearpage
\setstretch{1.3} % Reset the line-spacing to 1.3 for body text (if changed)
\acknowledgements{
    \addtocontents{toc}{\vspace{1em}} %Add a gap in the Contents, for aesthetic
    
    To my friends \textit{I Cazzilli}: Fede, Lorenz, Grismina, and Ste, for always being worthy of being the family I chose for myself and for looking out for me day after day.
    
    To Emme, for showing me what true compassion is and inspiring me to be a better version of myself every day.
    
    	To Carol, the closest tea-brewing sister I could hope for. I'm truly proud of you.
    
	To my Mom, for teaching me, among another thousand things, her own personal special way of calling somewhere \textit{``home"}.
	
	To my Dad, for always trying to be the man he ought to be.
	
	To the rest of my strange, enlarged family, for always supporting me in every possible way.
	
	To Luca, my tireless desk-mate and friend, who never called himself out of an all-nighter and who always speaks a word of kindness or dedication in times of need.
    
    To Valentina, who helped me stand in times when I couldn't stand for myself.
    
    To Matteo, an amazing companion in this journey through pediatrics, data sets, and enthusiastic English-vocabulary show off.
    
    To dott. Monasta, for making sense of numbers that didn't seem to have any.
    
    To prof. Barbi and prof. Ventura, for remembering me that medicine can be the marvelous cosmos I dreamed of. 
    
    And lastly, to myself, for always believing that, even when your heart's lost all its hope, after dawn there will be sunshine.
}