\bibliographystyle{unsrtnat}
\begin{thebibliography}{50}

\bibitem{nelson}
  Kliegman R, Stanton B, St Geme J, Schor N.
  \textit{Nelson: Textbook of Pediatrics}.
  20th edition, Vol. 1 and 2,
  pp. ~223-227.
  Elsevier (Philadelphia).
  2016.
  Available at: \url{https://www.elsevier.ca/ca/product.jsp?isbn=9781455775668} [accessed September 2018]
  
\bibitem{unreport}
  Department of Economic and Social Affairs of the United Nations Secretariat, Population Division.
  \textit{Child Adoption: Trends and Policies}.
  pp. ~45-48, 66-68, 74-83.
  New York,
  2009.
  Available at: \url{http://www.un.org/esa/population/publications/adoption2010/child_adoption.pdf} [accessed September 2018]

\bibitem{usreport}
  Bureau of Consular Affairs, U.S. Department of State.
  \textit{Annual Report on Intercountry Adoptions - FY 2017}.
  March 2008.
  Available at: \url{https://travel.state.gov/content/dam/NEWadoptionassets/pdfs/Annual%20Report%20on%20Intercountry%20Adoptions%20FY2017%20(release%20date%20March%2023%2020.._.pdf} [accessed September 2018]
  
\bibitem{usreportsite}
  Bureau of Consular Affairs, U.S. Department of State website.
  \textit{Adoption reference publications page}.
  Available at: \url{https://travel.state.gov/content/travel/en/Intercountry-Adoption/adopt_ref/adoption-publications.html} [accessed September 2018]

\bibitem{adoptdropping_article}
  Montgomery M, Powell I.
  \textit{International adoptions have dropped 72 percent since 2005 – here’s why}.
  The Conversation website.
  February 2018.
  Available at: \href{http://theconversation.com/international-adoptions-have-dropped-72-percent-since-2005-heres-why-91809}{\nolinkurl{http://theconversation.com/international-adoptions-have-dropped-72-percent-since-2005-heres-}
  \\
  \nolinkurl{why-91809}} [accessed September 2018]
  
\bibitem{adoptdropping_book}
  Montgomery M, Powell I.
  \textit{Saving International Adoption: An Argument from Economics and Personal Experience}.
  Vanderbilt University Press,
  pp. ~26-280.
  January 2018.
  
\bibitem{notonlyinfectious}
  Totaro C, Bortone B, Putignano P, Sollai S, Galli L, De Martino M, Chiappini E.
  \textit{Internationally adopted children: not only infectious diseases!}.
  Journal of Travel Medicine,
  Brief Communication,
  pp. ~1–5.
  2008.
  \doi{10.1093/jtm/tay036}
  
\bibitem{nonsoloinfezioni}
  Totaro C, Bortone B, Putignano P, Galli L, De Martino M, Chiappini E.
  \textit{Bambino adottato: non pensiamo solo alle infezioni!}.
  Medico e Bambino, Vol. 37, No. 5/2018,
  pp. ~307–3012.
  May 2018.
  Available at: \url{https://www.medicoebambino.com/?id=1805_307.pdf} [accessed September 2018]
  
\bibitem{cai2014}
  Italian Presidency of Council of Ministers, Commission for International Adoptions.
  \textit{Data e perspectives in international adoptions: years 2014-2015}.
  International Adoptions.
  pp. ~6-25.
  2016.
  Available at: \url{http://www.commissioneadozioni.it/media/153043/report_statistico_2014-2015.pdf} [accessed September 2018]  
  
\bibitem{cai2016}
  Italian Presidency of Council of Ministers, Commission for International Adoptions.
  \textit{Children and couples in international adoptions: years 2016-2017}.
  International Adoptions.
  pp. ~2-13.
  2018.
  Available at: \url{http://www.commissioneadozioni.it/media/155472/reportcai20162017.pdf} [accessed September 2018]  
  
\bibitem{caringfor}
  Miller L.
  \textit{Caring for Internationally Adopted Children}.
  New England Journal of Medicine. Vol 341, No. 20.
  Editorials,
  pp. ~1539–1545.
  November 1999.
  \doi{10.1056/NEJM199911113412009}
  
\bibitem{initialeval}
  American Academy of Pediatrics Committee on Early Childhood, Adoption and Dependent Care.
  \textit{Initial Medical Evaluation of an Adopted Child}.
  Pediatrics. Vol 88, No. 3.
  pp. ~642-644.
  September 1991.
  Available at: \url{http://pediatrics.aappublications.org/content/pediatrics/88/3/642.full.pdf} [accessed September 2018]
  
\bibitem{redbook}
  American Academy of Pediatrics Committee on Infectious Diseases.
  \textit{Red Book: 2012 Report of the Committee on Infectious Diseases}.
  \nth{29} Edition.
  pp. ~116-120.
  Elk Grove Village, IL.
  2012.

\bibitem{medicalscreening}
  Stauffer W, Maroushek S, and Deeak Kamat.
  \textit{Medical Screening of Immigrant Children}.
  Clinical Pediatrics, No. 42.
  pp. ~763-773.
  November 2003.
  \doi{10.1177/000992280304200902}
  
\bibitem{immunizationstatus}
  Viviano E, Cataldo F, Accomando S, Firenze A, Valenti RM, Romano N.
  \textit{Immunization status of internationally adopted children in Italy}.
  Vaccine, No. 24.
  pp. ~4138-4143.
  2006.
  \doi{10.1016/j.vaccine.2006.02.029}
  
\bibitem{comphealtheval}
  Jones V, and Committee on Early Childhood, Adoption and Dependent Care.
  \textit{Comprehensive Health Evaluation of the Newly Adopted Child}.
  Pediatrics, No. 129.
  Clinical Report,
  pp. ~e214-e223.
  2012.
  \doi{10.1542/peds.2011-2381}
  
\bibitem{immunizationstillinsufficient}
  van Schaik R, Wolfs T, Geelen S.
  \textit{Improved general health of international adoptees, but immunization status still insufficient}.
  European Journal of Pediatrics, No. 168.
  pp. ~1101-1106.
  2009.
  \doi{10.1007/s00431-008-0895-7}
  
\bibitem{circolareministeriale}
  Ministero della Salute, Governo della Repubblica Italiana.
  \textit{Circolare Ministeriale n.8 del 23/3/1993}.
  1993.

\bibitem{centroprevenzione}
  Ministero della Salute, Centro Nazionale Prevenzione e Controllo delle Malattie (CCM).  
  Available at: \url{http://www.ccm-network.it} [accessed September 2018]
  
\bibitem{GNLBI1}
  National Working Group for Migrant Children, Italian Society of Pediatrics.
  \textit{Standard Operating Procedure for Health Assistance of Newly Arrived Migrant Minors}.
  GLNBI-SIP Consensus Conference (Rome, 30/11/2013).
  Available at: \url{http://www.glnbi.org/documenti/7dc63f2c92f23828da9b9727d0a2779d.pdf} [accessed September 2018]
  
\bibitem{GNLBI2}
  Adami Lami C, Gabrielli C, Zaffaroni M, Cataldo F, Valentini P, Veneruso G, and Zavarise G.
  \textit{Nuovo protocollo diagnostico-assistenziale del GLNBI per il bambino adottato all'estero}.
  \nth{3} National Convention ISMU-GLNBI-SIMM (Milan, 17-18/05/2007).
  Available at: \url{http://www.glnbi.org/documenti/ad88407a4ca3bff42fa5666d67028df0.pdf} [accessed September 2018]
  
\bibitem{GNLBI3}
  National Working Group for Migrant Children, Italian Society of Pediatrics.
  \textit{Indicazioni per l’accoglienza sanitaria al minore migrante: il testo definitivo}.
  Pediatria, No. 11.
  November 2014.
  Available at: \url{http://www.glnbi.org/documenti/0f2233180e0e77187af0dd7e2230434b.pdf} [accessed September 2018]
  
\bibitem{GNLBIexp}
  Valentini P, Gargiullo L, Ceccarelli M, Ranno O.
  \textit{Health status of internationally adopted children. The experience of an Italian “GNLBI” Paediatric Centre}.
  Italian Journal of Public Health, Vol. 9, No. 3,
  pp. ~1-10.
  2012.
  \doi{10.2427/7527}
  
\bibitem{preventing}
  Chen L, Barnett E, and Wilson M.
  \textit{Preventing Infectious Diseases during and after International Adoption}.
  Annals of Internal Medicine, Vol. 139, No. 5, Part 1,
  Reviews,
  pp. ~371-379.
  2003.
  \doi{10.7326/0003-4819-139-5\_Part\_1-200309020-00013}
  
\bibitem{misdiagnosis}
  Chansoff IJ, Wells AM, and King L.
  \textit{Misdiagnosis and missed diagnoses in foster and adopted children with prenatal alcohol exposure}.
  Pediatrics, Vol. 135, No. 2.
  pp. ~264-270.
  February 2015.
  \doi{10.1542/peds.2014-2171}
  
\bibitem{sollai}
  Sollai S, Ghetti F, Bianchi L, De Martino M, Galli L, and Chiappini E.
  \textit{Infectious diseases prevalence, vaccination coverage, and diagnostic challenges in a population of internationally adopted children referred to a Tertiary Care Children’s Hospital from 2009 to 2015}.
  Medicine (Baltimore), Vol. 96, No. 12,
  Observational Study.
  pp. ~1-10.
  March 2017.
  \doi{10.1097/MD.0000000000006300}
  
\bibitem{ampofo}
  Ampofo K.
  \textit{Infectious disease issues in adoption of young children}.
  Current Opinion in Pediatrics, Vol. 25, No. 1.
  pp. ~78-87.
  February 2013.
  \doi{10.1097/MOP.0b013e32835c1357}
  
\bibitem{obringer}
  Obringer E, Walsh L.
  \textit{Infectious Diseases and Immunizations in International Adoption}.
  Pediatric Annals, Vol. 46, No. 2.
  pp. ~e56-e60.
  February 2017.
  \doi{10.3928/19382359-20170117-01}
  
\bibitem{vitDadopt1}
  Chiappini E, Vierucci F, Ghetti F, de Martino M, and Galli L.
  \textit{Vitamin D Status and Predictors of Hypovitaminosis D in Internationally Adopted Children}.
  PLoS ONE, Vol. 11, No. 9.
  pp. ~1-12.
  September 2016.
  \doi{10.1371/journal.pone.0158469}
  
\bibitem{vitDadopt2}
  Gustafson K, Eckerle J, Howard C, Andrews B, and Polgreen L.
  \textit{Prevalence of Vitamin D Deficiency in International Adoptees Within the First 6 Months After Adoption}.
  Clinical Pediatrics, Vol. 52, No. 12.
  pp. ~1149–1153.
  July 2013.
  \doi{10.1177/0009922813495955}
  
\bibitem{vitDadopt3}
  Raffaldi I, Garazzino S, Ballardini G , Zaffaroni M, Grasso N, Bona G, Tovo P, Guala A.
  \textit{Vitamin D Status in Internationally Adopted Children: the experience of Northwest Italy}.
  Minerva Pediatrica, No. 9.
  pp. ~1-9.
  September 2017.
  \doi{10.23736/S0026-4946.17.04883-6}
  
\bibitem{vitDadopt4}
  Franchi B, Piazza M, Sandri M, Tenero L, Comberiati P, Boner AL, Capristo C.
  \textit{25-hydroxyvitamin D serum level in children of different ethnicity living in Italy}.
  European Journal of Pediatrics, Vol. 174, Issue 6.
  pp. ~749–757.
  June 2015.
  \doi{10.1007/s00431-014-2451-y}
  
\bibitem{vitDcutoff1}
  Saggese G, Vierucci F, Prodam F, Cardinale F, Cetin I, Chiappini E, de’ Angelis GL, Massari M, Miraglia Del Giudice E, Miraglia Del Giudice M, Peroni D, Terracciano L, Agostiniani R, Careddu D, Ghiglioni DG, Bona G, Di Mauro G, and Corsello G.
  \textit{Vitamin D in pediatric age: consensus of the Italian Pediatric Society and the Italian Society of Preventive and Social Pediatrics, jointly with the Italian Federation of Pediatricians}.
  Italian Journal of Pediatrics.
  pp. ~44-51.
  December 2018.
  \doi{10.1186/s13052-018-0488-7}
  
\bibitem{vitDcutoff2}
  Allgrovea J, Shawb N.
  \textit{A Practical Approach to Vitamin D Deficiency and Rickets}.
  \nth{2} revised edition of \textit{Calcium and Bone Disorders in Children and Adolescents}, Chapter 8.
  Endocrine Development, Vol. 28.
  Basel, Karger.
  pp. ~119-133.
  2015.
  \doi{10.1159/000381000}
  
\bibitem{vitDcutoff3}
  Pearce S, Cheetham T.
  \textit{Diagnosis and management of vitamin D deficiency}.
  British Medical Journal, Vol. 340, Article b5664.
  pp. ~142-147.
  January 2010.
  \doi{10.1136/bmj.b5664}
  
\bibitem{Hbcutoff}
  World Health Organization (WHO).
  \textit{Haemoglobin concentrations for the diagnosis of anaemia and assessment of severity}.
  Vitamin and Mineral Nutrition Information System.
  pp. ~1-6.
  2011.
  \doi{10.1136/bmj.b5664}
  Available at: \url{http://www.who.int/vmnis/indicators/haemoglobin} [accessed September 2018]
  
\bibitem{MCVferritincutoff}
  Åsberg AE, Mikkelsen G, Aune M, Åsberg A.
  \textit{Empty iron stores in children and young adults - the diagnostic accuracy of MCV, MCH, and MCHC}.
  International Journal of Laboratory Hematology, Vol. 36, Issue 1.
  pp. ~98-104.
  February 2014.
  \doi{10.1111/ijlh.12132}
  
\bibitem{vitDcutoff_letter}
  Arundel P, Gemmell AS, Allgrove J, Bishop NJ, Burren CP, Jacobs B, Mughal MZ, Offiah AC, Shaw NJ.
  \textit{British Paediatric and Adolescent Bone Group’s position statement on vitamin D deficiency}.
  British Medical Journal, Vol. 345, Article e8182.
  p. 1.
  December 2012.
  \doi{10.1136/bmj.e8182}
  
\bibitem{height1}
  World Health Organization (WHO).
  \textit{WHO Child Growth Standards: Length/height-for-age, weight-for-age, weight-for-length, weight-for-height and body mass index-for-age. Methods and development}.
  pp. ~139-226.
  2006.
  Available at: \url{http://www.who.int/childgrowth/standards/velocity/tr3_velocity_report.pdf?ua=1} [accessed September 2018]
   
\bibitem{height2}
  World Health Organization (WHO).
  \textit{WHO Child Growth Standards: Growth velocity based on weight, length and head circumference. Methods and development}.
  pp. ~11-88.
  2009.
  Available at: \url{http://www.who.int/childgrowth/standards/velocity/tr3_velocity_report.pdf?ua=1} [accessed September 2018]
   
\bibitem{height3}
  De Wit C, Sas T, Wit J, and CutfieldW.
  \textit{Patterns of Catch-Up Growth}.
  The Journal of Pediatrics, Vol. 162, No. 2.
  Commentary, pp. ~415-420.
  October 2012.
  \doi{10.1016/j.jpeds.2012.10.014}
  
\bibitem{height4}
  Grellety E, and Golden M.
  \textit{Severely malnourished children with a low weight-for-height have a higher mortality than those with a low mid-upper-arm-circumference: Empirical data demonstrates Simpson’s paradox}.
  Nutrition Journal, Vol. 79, No. 17.
  pp. ~1-21.
  September 2018.
  \doi{10.1186/s12937-018-0384-4}
  
\bibitem{height5}
  Macdonald-Wallis C, Solomon-Moore E, Sebire S, Thompson J, Lawlor D, and Jago R.
  \textit{A longitudinal study of the associations of children’s body mass index and physical activity with blood pressure}.
  PLoS ONE, No. 12.
  pp. ~1-15.
  December 2017.
  \doi{10.1371/journal.pone.0188618}
  
\bibitem{nomenclatore}
  \textit{Nomenclatore tariffario di specialistica ambulatoriale}.
  Regione Autonoma Friuli-Venezia Giulia.
  May 2018.
  Available at: \url{http://www.regione.fvg.it/rafvg/cms/RAFVG/salute-sociale/sistema-sociale-sanitario/FOGLIA51/} [accessed September 2018]

\end{thebibliography}